---
title: "MA 5790 Project"
author: "Ezequiel Carrillo and Samantha Hair"
date: "1/22/2022"
output: html_document
---

```{r setup}
library(purrr)
library(tidyselect)
library(dplyr)
library(tidyr)
library(ggplot2)
library(dplyr)
library(Hmisc)
library(e1071)
library(caret)
library(corrplot)
library(moments)
library(kernlab)
```

# Data Description 
For this project, we are using the dataset 'Predicting Divorce' (https://www.kaggle.com/csafrit2/predicting-divorce). The goal of using this dataset, is to predict if a couple is likely to get divorced in the future. All 54 of the predictor variables are categorical and are ranked on a scale from 0-4, with 0 being the lowest and 4 being the highest. The variable that is the outcome, is Divorce (Y/N), coded in binary. The predictor variables as described on Kaggle are as follows: 

1.	If one of us apologizes when our discussion deteriorates, the discussion ends.
2.	I know we can ignore our differences, even if things get hard sometimes.
3.	When we need it, we can take our discussions with my spouse from the beginning and correct it.
4.	When I discuss with my spouse, to contact him will eventually work.
5.	The time I spent with my wife is special for us.
6.	We don't have time at home as partners.
7.	We are like two strangers who share the same environment at home rather than family.
8.	I enjoy our holidays with my wife.
9.	I enjoy traveling with my wife.
10.	Most of our goals are common to my spouse.
11.	I think that one day in the future, when I look back, I see that my spouse and I have been in harmony with each other.
12.	My spouse and I have similar values in terms of personal freedom.
13.	My spouse and I have similar sense of entertainment.
14.	Most of our goals for people (children, friends, etc.) are the same.
15.	Our dreams with my spouse are similar and harmonious.
16.	We're compatible with my spouse about what love should be.
17.	We share the same views about being happy in our life with my spouse
18.	My spouse and I have similar ideas about how marriage should be
19.	My spouse and I have similar ideas about how roles should be in marriage
20.	My spouse and I have similar values in trust.
21.	I know exactly what my wife likes.
22.	I know how my spouse wants to be taken care of when she/he sick.
23.	I know my spouse's favorite food.
24.	I can tell you what kind of stress my spouse is facing in her/his life.
25.	I have knowledge of my spouse's inner world.
26.	I know my spouse's basic anxieties.
27.	I know what my spouse's current sources of stress are.
28.	I know my spouse's hopes and wishes.
29.	I know my spouse very well.
30.	I know my spouse's friends and their social relationships.
31.	I feel aggressive when I argue with my spouse.
32.	When discussing with my spouse, I usually use expressions such as ‘you always’ or ‘you never’ .
33.	I can use negative statements about my spouse's personality during our discussions.
34.	I can use offensive expressions during our discussions.
35.	I can insult my spouse during our discussions.
36.	I can be humiliating when we discussions.
37.	My discussion with my spouse is not calm.
38.	I hate my spouse's way of open a subject.
39.	Our discussions often occur suddenly.
40.	We're just starting a discussion before I know what's going on.
41.	When I talk to my spouse about something, my calm suddenly breaks.
42.	When I argue with my spouse, ı only go out and I don't say a word.
43.	I mostly stay silent to calm the environment a little bit.
44.	Sometimes I think it's good for me to leave home for a while.
45.	I'd rather stay silent than discuss with my spouse.
46.	Even if I'm right in the discussion, I stay silent to hurt my spouse.
47.	When I discuss with my spouse, I stay silent because I am afraid of not being able to control my anger.
48.	I feel right in our discussions.
49.	I have nothing to do with what I've been accused of.
50.	I'm not actually the one who's guilty about what I'm accused of.
51.	I'm not the one who's wrong about problems at home.
52.	I wouldn't hesitate to tell my spouse about her/his inadequacy.
53.	When I discuss, I remind my spouse of her/his inadequacy.
54.	I'm not afraid to tell my spouse about her/his incompetence.

The variable that is the outcome, is Divorce (Y/N), coded in binary. 

```{r dataset}
data_raw <- read.csv("divorce.csv") 

data_raw <- data_raw %>% 
  mutate(Divorce_Y_N = as.factor(Divorce_Y_N))

data_predict <- data_raw[, -55]

dim(data_raw)
```

```{r explore the data}
#histograms for first 20 predictors
data_predict[,seq(1:20)] %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins = 9)

#histograms for predictors 21-40
data_predict[,seq(21:40)] %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins = 9)

#histograms for predictors 41-55
data_predict[,seq(41:55)] %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram(bins = 9)

#between predictor correlations
corr <- cor(data_predict[, -55])

corrplot(corr, order = "hclust")

#frequency table
freq_tab <- lapply(data_predict, table)

#missing data
sum(!complete.cases(freq_tab))

#calculate skewness for all predictors
predictorSkew=round(skewness(data_predict),3);

#plot skewness
boxplot(predictorSkew, main='Spread of Skewness for all predictors', xlab='predictors (54)', ylab='skewness coefficient');

plot(predictorSkew, main='scatter plot of predictor and its skew',xlab = 'predictor index',ylab = 'skewness coefficient')
hist(predictorSkew)
```

#Spending the data
```{r spending the data}
set.seed(1)

train_row <- createDataPartition(data_raw$Divorce_Y_N, p = .8, list = FALSE)

#train data
train_data <- data_raw[train_row, ]

train_predict <- train_data[, -55 ]

#test data
test_data <- data_raw[-train_row, ]
```

#Adding Predictors
Since all of our predictors are categorical, we do not need to add any predcitors. 

#Deleting Predictors
```{r deleting predictors}
novar <- nearZeroVar(train_predict)
corr <- cor(train_predict)

highcor <- findCorrelation(corr, cutoff = .90)

#remove predictors that are highly correlated
data_trans1 <- train_predict[,-highcor]
new_train <- train_data[,-highcor]

dim(data_trans1)
```


#Imputation
Since we do not have any missing values, we have no need to imputate any variables. 

#Center and Scale
No need to center and scale since all variables are on the same categorical scale of 0-4. 

#Spatial Sign
No need for spatial sign 

#Box-Cox transformation
#PC
```{r transforming the data}
trans <- preProcess(data_trans1, method = c("BoxCox","pca"))

#scree plot
results <- prcomp(data_trans1, center = TRUE, scale = TRUE)
var_explained <- results$sdev^2 / sum(results$sdev^2)

qplot(c(1:33), var_explained) + 
  geom_line() + 
  xlab("Principal Component") + 
  ylab("Variance Explained") +
  ggtitle("Scree Plot") +
  ylim(0, 1)


data_new <- predict(trans, data_trans1)


dim(data_new)
```


#Resampling the data
```{r resampling}
svm_fit <- train(Divorce_Y_N ~., data = new_train, method = "svmRadial", preProc = c("BoxCox", "pca"), tuneLength = 10, trControl = trainControl(method = "repeatedcv", repeats  = 5))

knn_fit <- train(Divorce_Y_N ~., data = new_train, method = "knn", preProc = c("BoxCox", "pca"), tuneLength = 5, trControl = trainControl(method = "repeatedcv", repeats  = 5))
```

#Finding the best resampling method
```{r}
resamp <- resamples(list(SVM = svm_fit, KNN = knn_fit))
summary(resamp)

model_diff <- diff(resamp)
summary(model_diff)
```

#Calculate R^2
```{r performance}
svm_pred <- predict(svm_fit, test_data[, -55])
svm_val <- data.frame(obs = test_data[, 55], pred = svm_pred)
defaultSummary(svm_val)
  
knn_pred <- predict(knn_fit, test_data[, -55])  
knn_val <- data.frame(obs = test_data[, 55], pred = knn_pred)
defaultSummary(knn_val)

```

